\section{Project Management}
% This section should clearly identify the work division among the students in
% the group using a project timeline that lists the tasks and milestones of the project, the member(s) of
% the group responsible for each task, and the actual completion date of each task. This section should
% contain a description of how the overall project developed and any successes and/or challenges during
% the course of the project.

The initial version of the project called for us to implement 4 algorithms: Elliptic Curve Cryptography, RSA, AES, and SHA. 
Alex was supposed to do elliptic curve, Jacob RSA, Drew AES, and Aaron SHA.
However, elliptic curve cryptography proved to be difficult to implement, so Alex joined forces with Aaron to work
on a Linux password cracker.
From that point on, we all worked individually on our own projects and got them working.
The following list details the work performed by each member:

\begin{itemize}
\item \textbf{Drew} (AES)

\begin{itemize}
\item Implement simple software AES (2018-11-07).
\item Synthesizable baseline AES (2018-11-15).
\item Optimizing AES by using apuints and rewriting (2018-11-20).
\item Optimized CTR AES complete (2018-11-29).
\end{itemize}

\item \textbf{Jacob} (RSA)
\begin{itemize}
\item Implement optimized software baseline RSA using GMP (2018-11-07).
\item Implemented very bad RSA baseline (2018-11-15).
\item Removed most variable bound loops from algorithm (2018-11-20).
\item Work on meeting timing closure and optimized version (2018-11-29).
\end{itemize}

\item \textbf{Aaron} (SHA)
\begin{itemize}
\item Optimized SHA256 for CPU (2018-11-07).
\item Synthesizable SHA256 (2018-11-07).
\item Synthesizable SHA512 (2018-11-20).
\item Created Unix password cracker (5000 rounds with special salting) (2018-11-29).
\end{itemize}

\item \textbf{Alex} (SHA)
\begin{itemize}
\item Investigate Elliptic Curves, decide on Secp256k1 curve (2018-11-07).
\item Abort ECC, switch to password cracker. Design simple brute force algorithm (2018-11-15).
\item Explore F1 (2018-11-29).
\item Write host program for password cracker (2018-11-29).
\end{itemize}
\end{itemize}
